---
title: "FML_Assignment1_Urooj"
output:
  html_document: default
  pdf_document: default
  word_document: default
date: "2023-09-08"
---

#1. The dataset used in this Assignment is PayEquity. 

#2.The data is imported from https://github.com/davidcaughlin/R-Tutorial-Data-Files/blob/master/PayEquity.csv 

```{r}
#install.packages("readr")    #"readr" package is installed to read the files
library(readr)                #library is used to load the package
setwd("C:\\Users\\user\\Desktop\\Masters\\Assignments\\FML Assignments\\Urooj_FML_Assignment")
read.csv(file="PayEquity.csv",header=TRUE,sep=",") #raw data file is imported using the                                                        read.csv function
Data<-read.csv(file="PayEquity.csv",header=TRUE,sep=",") #The imported file is assigned to                                                           a Data frame "Data"
View(Data)                    #View function is used to display the data
head(Data,5)                  # Used to find the 5 rows from data
```

#3.Descriptive statistics for a selection of quantitative and categorical variables

#3(a)Quantitative variable

```{r}
#install.packages("modeest")
library(modeest)         #estimating mode of distribution
summary(Data$job_tenure) #provides with min, max,mean,median and quartiles of attribute
mean(Data$job_tenure)    #calculates the average
median(Data$job_tenure)  #calculates the mid value
mfv(Data$job_tenure)     #most frequent value function is used to calculate mode
table(Data$job_tenure)   # table function shows all the values with its frequency
sd(Data$job_tenure)      # calculates the standard deviation
var(Data$job_tenure)     # calculates the Variance
length(Data$job_tenure)  # calculates the number of rows in column
quantile(Data$job_tenure)# calculates the quantile
range(Data$job_tenure)   # range calculates the min and max values
sum(Data$job_tenure)     # calculates the sum of column values
min(Data$job_tenure)     # calculates the minimum value in column
max(Data$job_tenure)     # calculates the maximum value in column
sort(Data$job_tenure)    # ascending order
sort(Data$job_tenure, decreasing=TRUE)# descending order
IQR(Data$job_tenure)     # calculates the InterQuartile Range
xtabs(~sex+race,data=Data)# Display value in cross tab
```

#3(b)Categorical variable

```{r}
table(Data$race)  # table function is used to count the frequency of each categorical value
prop.table(table(Data$race))*100 #calculates the proportion of different categories
mfv(Data$race)    # most frequent value function is used to find the highest frequent categorical value in a column
unique(Data$race) # Unique function is used to  find the unique values in a column by removing the duplicates
```

#4 Transforming a Quantitative Variable-Zscore, Square root

```{r}
#install.packages("ggplot2") # ggplot2 package is used to visualize data and plot the charts
library(ggplot2)            # ggplot2 is loaded to use for graphs
#install.packages("gridExtra")#gridExtra is used as an extension for grids
library(gridExtra)           # gridExtra is loaded for grids

#z-score
z=(Data$pay_grade_tenure-mean(Data$pay_grade_tenure))/sd(Data$pay_grade_tenure) #calculating the z score
z # The variable will give the output of zscore formula assigned to z 

#Square root
Sqrvar<-sqrt(Data$pay_grade_tenure) # sqrt function applies square root to each value in the column
Sqrvar # The variable will give the output of square root formula assigned to Sqrvar 

# to plot a histogram with pay_grade_tenure column ggplot is used and assigned to Var1 variable
Var1<- ggplot(data= Data, aes(x= pay_grade_tenure)) + geom_histogram()

# to plot a histogram with zscore (Z column) ggplot is used and assigned to Var2 variable
Var2<- ggplot(data= Data, aes(x= z)) + geom_histogram()

# to plot a histogram with square root (Sqrvar column) ggplot is used and assigned to Var3 variable
Var3<- ggplot(data= Data, aes(x= Sqrvar)) + geom_histogram()

# grid arrange is used to assemble the multiple plots in a page either in rows or columns specified
grid.arrange(Var1, Var2, Var3, ncol=1) 
```


#5.Plotting a quantitative variable, and one scatterplot
#5(a)Plotting a quantitative variable

```{r}
library(ggplot2)
ggplot(Data, aes(x = educ)) + geom_bar(stat = "count") + stat_count(geom = "text", aes(label = after_stat(count)), na.rm = TRUE) # a bar graph is plotted using ggplot function, x axis as "educ" column and y as count of "educ" by providing data values and labels for bars

ggplot(data = Data, aes(x = job_tenure, y = base_pay)) + geom_line()


```

#The above graph depicts the summary of base pay trends based on the job tenure. As the job tenure increases, the base pay trends are increasing.

#5.Plotting a Categorical variable
#5(b) pie chart

```{r}
#install.packages("plotrix") # plotrix is used to plot the pie chart
library(plotrix)            # package is loaded here
pie(table(Data$race), labels = names(table(Data$race)), col = rainbow(3)) # pie chart is drawn with race categorical value by providing names of data values and colouring pie
```

#The above piechart illustrates that there are major proportions of blacks, followed with white people. Asians are the lower population compare to other two groups.

#5(c) Scatter plot

```{r}
library(ggplot2) # loaded ggplot2 library to plot a scatterplot

ggplot(Data, aes(x = job_tenure, y = base_pay, color = race)) + geom_point() #scatter plot is plotted using job_tenure as x, base_pay as y and coloured data points from darker to lighter for higher to lower tenure

```

#The above graph illustrates that the majority of the data points lie between the tenure of 3-7 years of job tenure and their base pay is between 45,000 to 65,000.
#There is a high proportions of white and black followed with asian people.
#On the other hand, there are very few proportion of audience between 0-3 years and above 9 years and they are whites and blacks.
